Heavy rainfall in Paraguay during the 2015-2016 austral summer: causes and sub-seasonal-to-seasonal predictive skill

### 2019-01-27 reports/preprint/preprint.tex
add codes for preprint on eartharxiv

%-------------------------------------------------------------------------------
% Preamble
%-------------------------------------------------------------------------------
\documentclass{ametsoc}
\journal{jcli}

%Citations should be of the form ``author year''  not ``author, year''
\bibpunct{(}{)}{;}{a}{}{,}

%Packages
\usepackage{
	siunitx, % units
	cleveref,% cross-references
	physics, % better physics notation
	url, % for URLs
	amsmath, % for math
	array, % for fixed-width tables
	booktabs, % for nice tables
	nth, % for \nth{90}
}
\newcolumntype{L}[1]{>{\raggedright\let\newline\\\arraybackslash\hspace{0pt}}m{#1}}
\graphicspath{{../../figs/}{./}}

\usepackage[acronym]{glossaries}
\makeglossaries
\newacronym{wt}{WT}{weather type}
\newacronym{sallj}{SALLJ}{South American low-level jet}
\newacronym{enso}{ENSO}{El Ni\~{n}o-Southern Oscillation}
\newacronym{mjo}{MJO}{Madden-Julian Oscillation}
\newacronym{sacz}{SACZ}{South Atlantic Convergence Zone}
\newacronym{iri}{IRI}{International Research Institute for Climate and Society}
\newacronym{ecmwf}{ECMWF}{European Centre for Medium-Range Weather Forecasts}
\newacronym{s2s}{S2S}{sub-seasonal to seasonal}
\newacronym{eof}{EOF}{empirical orthogonal function}
\newacronym{mos}{MOS}{model output statistics}
\newacronym{xlr}{XLR}{homoscedastic extended logistic regression}
\newacronym{hxlr}{HXLR}{heteroscedastic extended logistic regression}
\newacronym{pcr}{PCR}{principal components regression}
\newacronym{cca}{CCA}{canonical correlation analysis}
\newacronym{sst}{SST}{sea-surface temperature}
\newacronym{scad}{SCAD}{South Central Atlantic Dipole}
\newacronym{lprb}{LPRB}{Lower Paraguay River Basin}

@@

\title{Heavy rainfall in Paraguay during the 2015-2016 austral summer: causes and sub-seasonal-to-seasonal predictive skill}
\authors{James Doss-Gollin\correspondingauthor{James Doss-Gollin. Columbia Water Center. Dept. of Earth and Environmental Engineering, Columbia University, 500 W. 120th St., New York, NY. USA.}}
\affiliation{Columbia Water Center. Dept. of Earth and Environmental Engineering, Columbia University, 500 W. 120th St., New York, NY. USA \\ Columbia Water Center, Columbia University, 500 W. 120th St., New York, NY. USA}
\email{james.doss-gollin@columbia.edu}
\extraauthor{\'Angel G. Mu\~noz}
\extraaffil{Atmospheric and Oceanic Sciences (AOS). Princeton University. Princeton, NJ 08540-6649. USA  \\
	International Research Institute for Climate and Society (IRI). The Earth Institute. Columbia University. Palisades, NY. 10964-1000. USA}
\extraauthor{Simon J. Mason}
\extraaffil{International Research Institute for Climate and Society (IRI). The Earth Institute. Columbia University. Palisades, NY. 10964-1000. USA}
\extraauthor{Max Past\'{e}n}
\extraaffil{Direcci\'{o}n de Meteorolog\'{i}a e Hidrolog\'{i}a. Paraguay \\ Facultad Polit\'ecnica. Universidad Nacional de Asunci\'{o}n. Paraguay}

%-------------------------------------------------------------------------------
% Abstract
% Up to 250 words in length
%-------------------------------------------------------------------------------

\abstract{
	During the austral summer 2015-16, severe flooding displaced over \num{170000} people on the Paraguay River system in Paraguay, Argentina, and Southern Brazil.
	These floods were driven by repeated heavy rainfall events in the Lower Paraguay River Basin.
	Alternating sequences of enhanced moisture inflow from the South American Low-Level Jet and local convergence associated with baroclinic systems were conducive to mesoscale convective activity and enhanced precipitation.
	These circulation patterns were favored by cross-timescale interactions of a very strong El Ni\~{n}o event, an unusually persistent Madden-Julian Oscillation in phases four and five, and the presence of a dipole SST anomaly in the central southern Atlantic Ocean.
	The simultaneous use of seasonal and sub-seasonal heavy rainfall predictions could have provided decision makers useful information about the start of these flooding events from two to four weeks in advance.
	Probabilistic seasonal forecasts available at the beginning of November successfully indicated heightened probability of heavy rainfall (\nth{90} percentile) over southern Paraguay and Brazil for December-February.
	Raw sub-seasonal forecasts of heavy rainfall exhibited limited skill at lead times beyond the first two predicted weeks, but a Model Output Statistics approach involving principal component regression substantially improved the spatial distribution of skill for week 3 relative to other methods tested including extended logistic regressions.
	A continuous monitoring of climate drivers impacting rainfall in the region, and the use of statistically corrected heavy precipitation seasonal and sub-seasonal forecasts, may help improve flood preparedness in this and other regions.
}

@@

@@

@@

@@

@@

@@

@@

@@

@@

@@

@@

@@

@@

@@

\acknowledgments
The authors thank David Farnham, Upmanu Lall, Laureline Josset, and Andrew Robertson for insightful conversations and guidance.
JDG thanks the NSF GRFP program for support (grant DGE 16-44869).
AGM was supported by the Atmospheric and Oceanic Sciences (AOS) Program at Princeton University.
Streamflow data shown in \cref{fig:streamflow} was collected by the Paraguayan Navy and National Administration of Navigation and Ports of Paraguay, and was processed and distributed by the Paraguayan Directorate of Meteorology and Hydrology (DINAC-DMH).
The authors also are thankful to the WWRP/WCRP Subseasonal-to-Seasonal Prediction Project Database, available via the IRI Data Library mirror at \url{https://iridl.ldeo.columbia.edu/SOURCES/.ECMWF/.S2S/}.

@@

\bibliographystyle{ametsoc2014}
\bibliography{../library}

%-------------------------------------------------------------------------------
% TABLES
%-------------------------------------------------------------------------------
\begin{table}
%
\caption{
	\Gls{mos} methods used to correct the \gls{ecmwf} sub-seasonal forecasts.
	Spatial domain for predictand is always the same (\SIrange{39}{17}{\degree S}; \SIrange{66}{49}{\degree W}).
	Two initializations are used: Nov 12th and 16th, 2015.
}\label{tab:mos-methods}
\begin{center}
	\begin{tabular}{L{0.75in} L{1.15in} L{3.5in}}
		\toprule
		Model & Region (Predictor) & Final predictor(s) selected \\
		%
		\midrule
		%
		\emph{Raw} & \SIrange{39}{17}{\degree S}; \SIrange{66}{49}{\degree W} & Ensemble mean, computed using members from the  two initializations. No correction performed. \\
		%
		\emph{XLR} & \SIrange{39}{17}{\degree S}; \SIrange{66}{49}{\degree W} & Ensemble mean, computed using members from the  two initializations.  \\
		%
		\emph{HLXR} & \SIrange{39}{17}{\degree S}; \SIrange{66}{49}{\degree W} & Ensemble mean and spread, computed using  members from the two initializations.\\
		%
		\emph{PCR} & \SIrange{60}{0}{\degree S}; \SIrange{80}{30}{\degree W} & Linear combination of model's \glspl{eof} computed using both initializations as independent predictors (10 \glspl{eof}).\\
		%
		\emph{CCA} & \SIrange{60}{0}{\degree S}; \SIrange{80}{30}{\degree W} & Canonical modes computed using both initializations as independent predictors. (10 predictor \glspl{eof}, 4 predictand \glspl{eof}, 4 canonical modes) \\
		%
		\bottomrule
		%
	\end{tabular}
\end{center}
\end{table}

%-------------------------------------------------------------------------------
% FIGURES
%-------------------------------------------------------------------------------

\begin{figure*}
	\noindent\includegraphics[width=6.5in]{study_area.jpg}
	\caption{
		Topographical map of the study area.
		Colors indicate $\log_{10}$ of elevation, in \si{\meter}, from the Global Land One-Km Base Elevation project available at \texttt{http://iridl.ldeo.columbia.edu/SOURCES/.NOAA/.NGDC/.GLOBE/.topo/}.
		(a): all of South America.
		The domains of the \acrlong{lprb} and the domain used for weather typing are indicated in red and blue, respectively.
		(b): The \acrfull{lprb}.
		As for (a), the \gls{lprb} is marked with a red box.
		Streamflow time series shown in \cref{fig:streamflow} were taken from the fours stations indicated.
		The Paraguay River and its tributaries, from the Natural Earth database (\texttt{www.naturalearthdata.com}), are also shown.
		Full station names are: Bah\'ia Negra (Bne); Concepci\'on (Conc); Asunci\'on (Asu); Pilar (Pil).
	}\label{fig:study-area}
\end{figure*}

\begin{figure*}
	\noindent\includegraphics[width=6.5in]{anomalies_ndjf1516.pdf}
	\caption{
		Monthly composite anomalies observed during NDJF 2015-16.
		Top row (a-d) shows streamfunction anomalies at \SI{850}{\hecto\pascal}.
		Bottom row (e-h) shows rainfall anomalies, in units of \si{\milli\meter\per\day}.
	}\label{fig:anomalies}
\end{figure*}

\begin{figure*}
	\noindent\includegraphics[width=6.5in]{Streamflow.pdf}
	\caption{
		River stage (height; in \si{\meter}) for the Paraguay River at four gauges along the Paraguay River.
		The station names are shortened versions of those shown in \cref{fig:study-area}.
		(a): Seasonality (orange) and time series of 2015-16 observations (black) at each stream gauge.
		Seasonality was fit using local polynomial regression as implemented in the \texttt{locfit} package in the \textbf{R} statistical programming environment \citep{Loader1999}.
		(b): Time series of daily stage measurements from 1929 to 2016 at each station.
	}\label{fig:streamflow}
\end{figure*}

\begin{figure*}	\noindent\includegraphics[width=6.5in]{lagged_rain.pdf}
	\caption{
		Composite anomalies associated with heavy rainfall (\nth{90} percentile exceedance of area-averaged rainfall in the \acrlong{lprb}).
		Lagged composites are shown, by column, for $t = $ \SIlist{-2;-1;0;1}{\day} relative to the date of heavy rainfall.
		Top row (a-d) shows composite streamfunction and wind anomalies at \SIlist{850}{\hecto\pascal}.
		Strongest 5\% of wind anomaly vectors between \SI{60}{\degree S} and \SI{10}{\degree N} (all longitudes) are also shown.
		Bottom row (e-h) shows composite rainfall anomalies, in units of \si{\milli\meter\per\day}.
	}\label{fig:lagged-rain}
\end{figure*}

\begin{figure*}
	\noindent\includegraphics[width=6.5in]{eof_loadings.pdf}
	\caption{
		Loadings of the four leading \glspl{eof} of daily NDJF \SI{850}{\hecto\pascal} streamfunction over the weather typing region shown in \cref{fig:study-area}.
		Parentheses in sub-plot titles indicate the percentage of total variance explained by each \gls{eof}.
	}\label{fig:eof-loading}
\end{figure*}

\begin{figure*}
	\noindent\includegraphics[width=6.5in,height=9in,keepaspectratio=true]{wt_composite.pdf}
	\caption{
		Composite anomalies associated with each weather type.
		Top row (a-f) shows streamfunction anomalies at \SIlist{850}{\hecto\pascal}.
		Strongest 20\% of wind anomaly vectors over the plot area are also shown.
		Bottom row (g-l) shows rainfall anomalies, in units of \si{\milli\meter\per\day}.
		The relative frequency of occurrence of each weather type (in days) is presented on the top of each column.
	}\label{fig:wt-composite}
\end{figure*}

\begin{figure*}
	\noindent\includegraphics[width=6.5in]{rain_wt_201516.pdf}
	\caption{
		Time series of area-averaged rainfall in the \acrlong{lprb} (\cref{fig:study-area}) for each day of NDJF  2015-16.
		Lines indicate the rainfall value, in units of \si{\milli\meter\per\day}.
        The weather type corresponding to each day is indicated in an adjacent text label.
		Dashed lines blue indicate (from bottom to top) the climatological 50th, \nth{90}, and 99th percentiles of NDJF area-averaged rain over the \acrlong{lprb}.
	}\label{fig:rain-wt}
\end{figure*}

\begin{figure*}
	\noindent\includegraphics[width=3.5in]{seasonal_forecast.pdf}
	\caption{
		Seasonal model forecast for probability of exceedance of \nth{90} percentile of DJF rainfall, as issued in November 2015.
		Color indicates the forecast probability of exceeding the \nth{90} percentile of climatological rainfall during DJF 2015-16 -- this is presented as the odds ratio as defined in \cref{eq:odds-ratio}.
		A value greater than 1 indicates that the model forecast greater-than-average odds of rainfall exceeding the \nth{90} percentile.
		Grid cells which observed an exceedance of the \nth{90} percentile of DJF rainfall are outlined in black.
	}\label{fig:seas-prob-fcst}
\end{figure*}

\begin{figure*}
	\noindent\includegraphics[width=6.5in]{chiclet.pdf}
	\caption{
		Chiclet diagram \citep[see][]{Carbin2016} of ensemble-mean precipitation anomaly forecasts over the \acrlong{lprb} (see \cref{fig:study-area}) from uncorrected \gls{ecmwf} \gls{s2s} model forecast data, as a function of the forecast target date (horizontal axis) and lead time (vertical axis).
		Time series of CPC daily mean precipitation over the same area is plotted with $y$-axis inverted; horizontal black line denotes NDJF climatology.
	}\label{fig:chiclet}
\end{figure*}

\begin{figure*}
	\noindent\includegraphics[width=6.4in]{mos_forecasts.pdf}
	\caption{
		Raw and \gls{mos}-adjusted \gls{s2s} model forecasts and skill scores for the methods indicated in \cref{tab:mos-methods}.
		Top row (a-e) shows the heavy rainfall forecast for 1-7 December 2015 as the odds ratio defined in \cref{eq:odds-ratio} over the target domain.
		A value greater than 1 indicates that the model forecast greater-than-average odds of rainfall exceeding the \nth{90} percentile.
		Second row (f-j) shows the Ignorance Score defined in \cref{eq:ignorance}, with zero indicating a perfect forecast.
		Bottom row (k-o) shows the 2AFC skill score for each grid cell; a value greater than 50 indicates that the model outperforms climatology.
		Columns separate different \gls{mos} models except for ``Raw'' (a,f,k), which indicates the uncorrected \gls{s2s} model output.
		For all rows the grid cells which observed a \nth{90} percentile exceedance for 1-7 December 2015 are outlined in black.
	}\label{fig:subs-prob-fcst}
\end{figure*}

\begin{figure*}
	\includegraphics[width=\textwidth]{wt_mjo_enso.pdf}
	\caption{
		Anomalous probability of occurrence of each weather type concurrent with observance of each \gls{mjo} phase.
		When \gls{mjo} amplitude is less than 1, it is defined as neutral phase (0).
		Plots are shown separately for El Ni\~no (NINO 3.4 $>1$), La Ni\~na (NINO 3.4 $< -1$), and Neutral \gls{enso} phases.
		Only values which are significant at $\alpha=0.10$, calculated with a bootstrap of 5000 samples, are shown.
	}\label{fig:wt-mjo-enso}
\end{figure*}

\begin{figure*}
	\noindent\includegraphics[width=6.4in]{ChacoNoChacojet.pdf}
	\caption{
		Schematics of low-level jet events (red arrows) during austral summer and El Ni\~no years.
		Most jet events are of the ``Chaco'' type, particularly when SST anomalies in the central southern Atlantic Ocean (a, see green box) are weak.
		When a dipole SST anomaly occurs in the central southern Atlantic with the warmer pole equatorward, the meridional temperature gradient and sea-land temperature contrasts establish an anticyclonic circulation (dot-dashed line) conducive to increased occurrence of No-Chaco jet events (b).
		Other SST anomaly configurations tend to be present outside the green box (not shown).
		Winds in panels are typical for each case (at \SI{850}{\hecto\pascal}).
		Reference wind vector in \si{\meter\per\second}.
		Green box shows location of \gls{scad}.
	}\label{fig:chaco-nochaco}
\end{figure*}

>>>>

\documentclass[12pt,letterpaper]{article}

% Page setup
\usepackage[
    margin=1in
]{geometry}

%Packages
\usepackage{
	amsmath, % for math
	array, % for fixed-width tables
	authblk, % for authors
	booktabs, % for nice tables
	graphicx, % figures
	natbib, % references
	nth, % for \nth{90}
	physics, % better physics notation
	siunitx, % units
	url, % for URLs
}
\newcolumntype{L}[1]{>{\raggedright\let\newline\\\arraybackslash\hspace{0pt}}m{#1}}
\graphicspath{{../../figs/}{./}}

% better looking captions
\usepackage[
  format=plain,
  labelfont={bf,it,footnotesize},
  textfont={it,footnotesize}
]{caption}

\usepackage[acronym]{glossaries}
\makeglossaries
\newacronym{cca}{CCA}{canonical correlation analysis}
\newacronym{ecmwf}{ECMWF}{European Centre for Medium-Range Weather Forecasts}
\newacronym{enso}{ENSO}{El Ni\~{n}o-Southern Oscillation}
\newacronym{eof}{EOF}{empirical orthogonal function}
\newacronym{iri}{IRI}{International Research Institute for Climate and Society}
\newacronym{hxlr}{HXLR}{heteroscedastic extended logistic regression}
\newacronym{lprb}{LPRB}{Lower Paraguay River Basin}
\newacronym{mjo}{MJO}{Madden-Julian Oscillation}
\newacronym{mos}{MOS}{model output statistics}
\newacronym{sallj}{SALLJ}{South American low-level jet}
\newacronym{pcr}{PCR}{principal components regression}
\newacronym{s2s}{S2S}{sub-seasonal to seasonal}
\newacronym{sacz}{SACZ}{South Atlantic Convergence Zone}
\newacronym{scad}{SCAD}{South Central Atlantic Dipole}
\newacronym{sst}{SST}{sea-surface temperature}
\newacronym{xlr}{XLR}{homoscedastic extended logistic regression}
\newacronym{wt}{WT}{weather type}

@@

\date{}
\title{Heavy rainfall in Paraguay during the 2015-2016 austral summer: causes and sub-seasonal-to-seasonal predictive skill}
\author[1,2]{James Doss-Gollin}
\author[3,4]{\'{A}ngel G. Mu\~{n}oz}
\author[4]{Simon J. Mason}
\author[5,6]{Max Past\'{e}n}
\affil[1]{Department of Earth and Environmental Engineering, Columbia University}
\affil[2]{Columbia Water Center}
\affil[3]{Atmospheric and Oceanic Sciences (AOS), Princeton University}
\affil[4]{International Research Institute for Climate and Society (IRI), The Earth Institute, Columbia University.}
\affil[5]{Direcci\'{o}n de Meteorolog\'{i}a e Hidrolog\'{i}a, Paraguay}
\affil[6]{Facultad Polit\'ecnica, Universidad Nacional de Asunci\'{o}n, Paraguay}

% These go last!
\usepackage[hidelinks]{hyperref}
\usepackage{cleveref}

@@

This is a pre-print of the paper ``Heavy rainfall in Paraguay during the 2015-16 austral summer: causes and sub-seasonal-to-seasonal predictive skill'', by James Doss-Gollin, \'{A}ngel G. Mu\~{n}oz, Simon J. Mason, and Max Past\'{e}n, published in Journal of Climate.
You can find the final, type-set version at \url{https://journals.ametsoc.org/doi/abs/10.1175/JCLI-D-17-0805.1} or see the full citation reference in the bibliography under \citet{DossGollin:2018bn}.

Codes used to build this paper are available at \url{https://github.com/jdossgollin/2018-paraguay-floods}.

\clearpage

@@

\begin{abstract}
	During the austral summer 2015-16, severe flooding displaced over \num{170000} people on the Paraguay River system in Paraguay, Argentina, and Southern Brazil.
	These floods were driven by repeated heavy rainfall events in the Lower Paraguay River Basin.
	Alternating sequences of enhanced moisture inflow from the South American Low-Level Jet and local convergence associated with baroclinic systems were conducive to mesoscale convective activity and enhanced precipitation.
	These circulation patterns were favored by cross-timescale interactions of a very strong El Ni\~{n}o event, an unusually persistent Madden-Julian Oscillation in phases four and five, and the presence of a dipole SST anomaly in the central southern Atlantic Ocean.
	The simultaneous use of seasonal and sub-seasonal heavy rainfall predictions could have provided decision makers useful information about the start of these flooding events from two to four weeks in advance.
	Probabilistic seasonal forecasts available at the beginning of November successfully indicated heightened probability of heavy rainfall (\nth{90} percentile) over southern Paraguay and Brazil for December-February.
	Raw sub-seasonal forecasts of heavy rainfall exhibited limited skill at lead times beyond the first two predicted weeks, but a Model Output Statistics approach involving principal component regression substantially improved the spatial distribution of skill for week 3 relative to other methods tested including extended logistic regressions.
	A continuous monitoring of climate drivers impacting rainfall in the region, and the use of statistically corrected heavy precipitation seasonal and sub-seasonal forecasts, may help improve flood preparedness in this and other regions.
\end{abstract}

@@

\begin{figure}
	\centering
	\includegraphics[width=6.5in]{study_area.jpg}
	\caption{
		Topographical map of the study area.
		Colors indicate $\log_{10}$ of elevation, in \si{\meter}, from the Global Land One-Km Base Elevation project available at \texttt{http://iridl.ldeo.columbia.edu/SOURCES/.NOAA/.NGDC/.GLOBE/.topo/}.
		(a): all of South America.
		The domains of the \acrlong{lprb} and the domain used for weather typing are indicated in red and blue, respectively.
		(b): The \acrfull{lprb}.
		As for (a), the \gls{lprb} is marked with a red box.
		Streamflow time series shown in \cref{fig:streamflow} were taken from the fours stations indicated.
		The Paraguay River and its tributaries, from the Natural Earth database (\texttt{www.naturalearthdata.com}), are also shown.
		Full station names are: Bah\'ia Negra (Bne); Concepci\'on (Conc); Asunci\'on (Asu); Pilar (Pil).
	}\label{fig:study-area}
\end{figure}
\begin{figure}
	\centering
	\includegraphics[width=6.5in]{anomalies_ndjf1516.pdf}
	\caption{
		Monthly composite anomalies observed during NDJF 2015-16.
		Top row (a-d) shows streamfunction anomalies at \SI{850}{\hecto\pascal}.
		Bottom row (e-h) shows rainfall anomalies, in units of \si{\milli\meter\per\day}.
	}\label{fig:anomalies}
\end{figure}

@@

\begin{table}
	\centering
	\caption{
		\Gls{mos} methods used to correct the \gls{ecmwf} sub-seasonal forecasts.
		Spatial domain for predictand is always the same (\SIrange{39}{17}{\degree S}; \SIrange{66}{49}{\degree W}).
		Two initializations are used: Nov 12th and 16th, 2015.
	}\label{tab:mos-methods}
	\begin{center}
		\begin{tabular}{L{0.75in} L{1.15in} L{3.5in}}
			\toprule
			Model & Region (Predictor) & Final predictor(s) selected \\
			%
			\midrule
			%
			\emph{Raw} & \SIrange{39}{17}{\degree S}; \SIrange{66}{49}{\degree W} & Ensemble mean, computed using members from the  two initializations. No correction performed. \\
			%
			\emph{XLR} & \SIrange{39}{17}{\degree S}; \SIrange{66}{49}{\degree W} & Ensemble mean, computed using members from the  two initializations.  \\
			%
			\emph{HLXR} & \SIrange{39}{17}{\degree S}; \SIrange{66}{49}{\degree W} & Ensemble mean and spread, computed using  members from the two initializations.\\
			%
			\emph{PCR} & \SIrange{60}{0}{\degree S}; \SIrange{80}{30}{\degree W} & Linear combination of model's \glspl{eof} computed using both initializations as independent predictors (10 \glspl{eof}).\\
			%
			\emph{CCA} & \SIrange{60}{0}{\degree S}; \SIrange{80}{30}{\degree W} & Canonical modes computed using both initializations as independent predictors. (10 predictor \glspl{eof}, 4 predictand \glspl{eof}, 4 canonical modes) \\
			%
			\bottomrule
			%
		\end{tabular}
	\end{center}
\end{table}

@@

\begin{figure}
	\centering
	\includegraphics[width=6.5in]{Streamflow.pdf}
	\caption{
		River stage (height; in \si{\meter}) for the Paraguay River at four gauges along the Paraguay River.
		The station names are shortened versions of those shown in \cref{fig:study-area}.
		(a): Seasonality (orange) and time series of 2015-16 observations (black) at each stream gauge.
		Seasonality was fit using local polynomial regression as implemented in the \texttt{locfit} package in the \textbf{R} statistical programming environment \citep{Loader1999}.
		(b): Time series of daily stage measurements from 1929 to 2016 at each station.
	}\label{fig:streamflow}
\end{figure}

@@

\begin{figure}
	\centering
	\includegraphics[width=6.5in]{lagged_rain.pdf}
	\caption{
		Composite anomalies associated with heavy rainfall (\nth{90} percentile exceedance of area-averaged rainfall in the \acrlong{lprb}).
		Lagged composites are shown, by column, for $t = $ \SIlist{-2;-1;0;1}{\day} relative to the date of heavy rainfall.
		Top row (a-d) shows composite streamfunction and wind anomalies at \SIlist{850}{\hecto\pascal}.
		Strongest 5\% of wind anomaly vectors between \SI{60}{\degree S} and \SI{10}{\degree N} (all longitudes) are also shown.
		Bottom row (e-h) shows composite rainfall anomalies, in units of \si{\milli\meter\per\day}.
	}\label{fig:lagged-rain}
\end{figure}

@@

\begin{figure}
	\centering
	\includegraphics[width=6.5in]{eof_loadings.pdf}
	\caption{
		Loadings of the four leading \glspl{eof} of daily NDJF \SI{850}{\hecto\pascal} streamfunction over the weather typing region shown in \cref{fig:study-area}.
		Parentheses in sub-plot titles indicate the percentage of total variance explained by each \gls{eof}.
	}\label{fig:eof-loading}
\end{figure}

\begin{figure}
	\centering
	\includegraphics[width=6.5in,height=9in,keepaspectratio=true]{wt_composite.pdf}
	\caption{
		Composite anomalies associated with each weather type.
		Top row (a-f) shows streamfunction anomalies at \SIlist{850}{\hecto\pascal}.
		Strongest 20\% of wind anomaly vectors over the plot area are also shown.
		Bottom row (g-l) shows rainfall anomalies, in units of \si{\milli\meter\per\day}.
		The relative frequency of occurrence of each weather type (in days) is presented on the top of each column.
	}\label{fig:wt-composite}
\end{figure}

@@

\begin{figure}
	\centering
	\includegraphics[width=6.5in]{rain_wt_201516.pdf}
	\caption{
		Time series of area-averaged rainfall in the \acrlong{lprb} (\cref{fig:study-area}) for each day of NDJF  2015-16.
		Lines indicate the rainfall value, in units of \si{\milli\meter\per\day}.
        The weather type corresponding to each day is indicated in an adjacent text label.
		Dashed lines blue indicate (from bottom to top) the climatological 50th, \nth{90}, and 99th percentiles of NDJF area-averaged rain over the \acrlong{lprb}.
	}\label{fig:rain-wt}
\end{figure}

@@

\begin{figure}
	\centering
	\includegraphics[width=3.5in]{seasonal_forecast.pdf}
	\caption{
		Seasonal model forecast for probability of exceedance of \nth{90} percentile of DJF rainfall, as issued in November 2015.
		Color indicates the forecast probability of exceeding the \nth{90} percentile of climatological rainfall during DJF 2015-16 -- this is presented as the odds ratio as defined in \cref{eq:odds-ratio}.
		A value greater than 1 indicates that the model forecast greater-than-average odds of rainfall exceeding the \nth{90} percentile.
		Grid cells which observed an exceedance of the \nth{90} percentile of DJF rainfall are outlined in black.
	}\label{fig:seas-prob-fcst}
\end{figure}

@@

\begin{figure}
	\centering
	\includegraphics[width=6.5in]{chiclet.pdf}
	\caption{
		Chiclet diagram \citep[see][]{Carbin2016} of ensemble-mean precipitation anomaly forecasts over the \acrlong{lprb} (see \cref{fig:study-area}) from uncorrected \gls{ecmwf} \gls{s2s} model forecast data, as a function of the forecast target date (horizontal axis) and lead time (vertical axis).
		Time series of CPC daily mean precipitation over the same area is plotted with $y$-axis inverted; horizontal black line denotes NDJF climatology.
	}\label{fig:chiclet}
\end{figure}

@@

\begin{figure}
	\centering
	\includegraphics[width=6.4in]{mos_forecasts.pdf}
	\caption{
		Raw and \gls{mos}-adjusted \gls{s2s} model forecasts and skill scores for the methods indicated in \cref{tab:mos-methods}.
		Top row (a-e) shows the heavy rainfall forecast for 1-7 December 2015 as the odds ratio defined in \cref{eq:odds-ratio} over the target domain.
		A value greater than 1 indicates that the model forecast greater-than-average odds of rainfall exceeding the \nth{90} percentile.
		Second row (f-j) shows the Ignorance Score defined in \cref{eq:ignorance}, with zero indicating a perfect forecast.
		Bottom row (k-o) shows the 2AFC skill score for each grid cell; a value greater than 50 indicates that the model outperforms climatology.
		Columns separate different \gls{mos} models except for ``Raw'' (a,f,k), which indicates the uncorrected \gls{s2s} model output.
		For all rows the grid cells which observed a \nth{90} percentile exceedance for 1-7 December 2015 are outlined in black.
	}\label{fig:subs-prob-fcst}
\end{figure}

@@

\begin{figure}
	\centering
	\includegraphics[width=\textwidth]{wt_mjo_enso.pdf}
	\caption{
		Anomalous probability of occurrence of each weather type concurrent with observance of each \gls{mjo} phase.
		When \gls{mjo} amplitude is less than 1, it is defined as neutral phase (0).
		Plots are shown separately for El Ni\~no (NINO 3.4 $>1$), La Ni\~na (NINO 3.4 $< -1$), and Neutral \gls{enso} phases.
		Only values which are significant at $\alpha=0.10$, calculated with a bootstrap of 5000 samples, are shown.
	}\label{fig:wt-mjo-enso}
\end{figure}

@@

\begin{figure}
	\centering
	\includegraphics[width=6.4in]{ChacoNoChacojet.pdf}
	\caption{
		Schematics of low-level jet events (red arrows) during austral summer and El Ni\~no years.
		Most jet events are of the ``Chaco'' type, particularly when SST anomalies in the central southern Atlantic Ocean (a, see green box) are weak.
		When a dipole SST anomaly occurs in the central southern Atlantic with the warmer pole equatorward, the meridional temperature gradient and sea-land temperature contrasts establish an anticyclonic circulation (dot-dashed line) conducive to increased occurrence of No-Chaco jet events (b).
		Other SST anomaly configurations tend to be present outside the green box (not shown).
		Winds in panels are typical for each case (at \SI{850}{\hecto\pascal}).
		Reference wind vector in \si{\meter\per\second}.
		Green box shows location of \gls{scad}.
	}\label{fig:chaco-nochaco}
\end{figure}

@@

\section*{Acknowledgments}
The authors thank David Farnham, Upmanu Lall, Laureline Josset, and Andrew Robertson for insightful conversations and guidance.
JDG thanks the NSF GRFP program for support (grant DGE 16-44869).
AGM was supported by the Atmospheric and Oceanic Sciences (AOS) Program at Princeton University.
Streamflow data shown in \cref{fig:streamflow} was collected by the Paraguayan Navy and National Administration of Navigation and Ports of Paraguay, and was processed and distributed by the Paraguayan Directorate of Meteorology and Hydrology (DINAC-DMH).
The authors also are thankful to the WWRP/WCRP Subseasonal-to-Seasonal Prediction Project Database, available via the IRI Data Library mirror at \url{https://iridl.ldeo.columbia.edu/SOURCES/.ECMWF/.S2S/}.

@@

\setglossarystyle{list}
\printglossary[type=\acronymtype]
\bibliographystyle{ametsoc2014}
\bibliography{../library}
